Parallel Information Set Generation for Kriegspiel

### 2012-07-16 sbac/paper.tex
more

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from White's perspective is shown in Figure~\ref{filteredlisting}.  Note that Black's actual moves have been
replaced with ``??'' and lists of Black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, Black attempted three illegal moves.
\begin{figure}
%\begin{minipage}[b]{0.35\linewidth}
%\centering
%\small
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
%\end{verbatim}
%\end{minipage}
%\begin{minipage}[b]{0.25\linewidth}
%\small
%\centering
%\begin{verbatim}
%\end{minipage}
%\vspace{-0.1in}
\caption{Observations received by White over the course of the game.  White does not know Black's actual moves but hears moderator announcements related to captures, pawn tries, and the number of attempted moves by Black that fail.}
\label{filteredlisting}
%\vspace{-0.2in}
\end{figure}

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\caption{Abbreviated output from example game.  Three of 25 sequences that are consistent with Black's observations after five moves by each player.  The 25 sequences (or the states that result from them) compose Black's information set at that point in the game. From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\end{figure}

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Core of Information Set Generation Algorithm}
\label{codelisting1}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our parallel ISG search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

Charm++ (and hence {\sc ParSSSE}) provides multiple load balancing strategies and we used three of them in our application. We describe these strategies next. The
first is the randomized strategy.  In this approach, each newly created object
is assigned to a random processor.  Generally, if a large enough number of 
chares is created, the system will achieve an approximately uniform
distribution of work across processors.  The approach is also appealing because
of its simplicity. However, randomization can sometimes lead to inefficiencies
from parallel overhead.  For a program running on $p$ processors, the
probability that a chare is assigned to a processor other than the
one that created it is $1-1/p$.  As the number of processors gets large, this
probability approaches 1, which means that the parallel overhead related to the
creation and scheduling of a new task is incurred for almost every piece of new
work.

@@

We also evaluated our information set generation routine on 40 positions from
actual Kriegspiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used Charm++'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.

@@

@@

\bibliographystyle{IEEEtran}
\bibliography{paper,cited}

>>>>

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from White's perspective is shown in Figure~\ref{filteredlisting}.  Note that Black's actual moves have been
replaced with ``??'' and lists of Black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, Black attempted three illegal moves.
\begin{figure}
%\begin{minipage}[b]{0.35\linewidth}
%\centering
%\small
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
%\end{verbatim}
%\end{minipage}
%\begin{minipage}[b]{0.25\linewidth}
%\small
%\centering
%\begin{verbatim}
%\end{minipage}
%\vspace{-0.1in}
\caption{Observations received by White over the course of the game.  White does not know Black's actual moves but hears moderator announcements related to captures, pawn tries, and the number of attempted moves by Black that fail.}
\label{filteredlisting}
\vspace{-0.2in}
\end{figure}

@@

%1. Pd2:d4 Pa7:a5 2. Bc1:f4 Pb7:b6 3. Bf4:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pa2:a4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pb2:b3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pc2:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pe2:e4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pf2:f4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Pg2:g4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ph2:h4 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:a3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:c3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Nb1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:d3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Qd1:c1 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ke1:d2 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:f3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
%1. Pd2:d4 Pa7:a5 2. Bc1:g5 Pb7:b6 3. Ng1:h3 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6 \\
\end{mytinylisting}
\caption{Abbreviated output from example game.  Three of 25 sequences that are consistent with Black's observations after five moves by each player.  The 25 sequences (or the states that result from them) compose Black's information set at that point in the game. From this, Black can infer that there is definitely a white bishop at g5 and a white pawn at d6.}
\label{abbrevoutput}
\vspace{-0.1in}
\end{figure}

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Core of Information Set Generation Algorithm}
\label{codelisting1}
\vspace{-0.2in}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our parallel ISG search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}

@@

{\sc Charm++} (and hence {\sc ParSSSE}) provides multiple load balancing strategies and we used three of them in our application~\cite{GengbinThesis}. We describe these strategies next. The
first is the randomized strategy.  In this approach, each newly created object
is assigned to a random processor.  Generally, if a large enough number of 
chares is created, the system will achieve an approximately uniform
distribution of work across processors.  The approach is also appealing because
of its simplicity. However, randomization can sometimes lead to inefficiencies
from parallel overhead.  For a program running on $p$ processors, the
probability that a chare is assigned to a processor other than the
one that created it is $1-1/p$.  As the number of processors gets large, this
probability approaches 1, which means that the parallel overhead related to the
creation and scheduling of a new task is incurred for almost every piece of new
work.

@@

We also evaluated our information set generation routine on 40 positions from
actual Kriegspiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used {\sc Charm++}'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.

@@

The effect of load balancing and computational granularity on combinatorial search problems has been studied before~\cite{sun11adaptive,WeatherAmpiSBAC10}. In this work, we addressed the parallelization of an information set generation algorithm and studied the impact of these factors applied to game tree applications. %We explored an application  which has largely been unexplored by the HPC community

@@

\bibliographystyle{IEEEtran}
\bibliography{paper,cited,group}
